###Bulk downloads of space and time targeted MODIS data
First thing you need to get the appropriate software installed, the [modis reprojection tool](https://lpdaac.usgs.gov/tools/modis_reprojection_tool) is used by this package to process your data. It can be used outside of R too, but this automates the process.

Then load packages, and set up your login and file path, followed by checking out what modis products are avaialble, and their appropriate codes. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      results = "hide")
```
```{r, results="markup"}
library(raster)
library(rgeos)
library(rgdal)
library(rts) #install and load the rts package for modistools specifically
#need to set nasa authorization (username and password that are separately acquired online, see link above), but only need to do once with the function below: setNASAauth()
#setNASAauth()
#setMRTpath("/Users/Maxwell/MRT/bin/", update=T) #function to set the path of the modis reprojection tool for the functions to find
#modisProducts() #function to check out possible products (Beware, the version of Modis matters for extracting with ModisDownload later)
```

```{r, results="markup"}
#make a date sequence to extract
start<-NULL
end<-NULL
start<-paste0(rep(2016:2018,1) , rep(".01.01",2))
end<-paste0(rep(2016:2018,1),rep(".05.31",2))
start
end
#combine into one statement for each desired date so it is easy to loop
dateseq<-NULL
dateseq<-paste((start), noquote("\",\""), paste(end), sep = "") 
dateseq
```

Now to extract dates we have set up from a particular modis product
```{r}
#extract dateseq for all of oregon, mosaic, and reproject all in one step using ModisDownload
  ##note the "MOD13A3" is the modis product
    ###h and v are the modis tiles (can look up online which oncs youll need)
      ####bands - each modis product has a bunch of bands, so here can specify which ones you want, need to put 0 or 1 for all or else it may do somethin weird
        #####proj_params Something weird, look up in the package to get a better idea of what to do here.
          ######utm_zone only needed if using UTM
            ######Datum: projection, pixel size: set what you what, mosaic: default is FALSE, need to set to true to mosaic. If you want there is a separate funciton to do it manually. Same                with proj, which reprojects the mosaicked file to a tif. Versopm: important, seems like 006 has everything I wanted, 005 didnt have much.
#download all modis files in the date range to your working directory (set above)

ModisDownload("MOD13A3",h=c(8, 9),v=c(4,4),dates=c("2016.01.01", "2016.05.31"),bands_subset="1 0 0 0 0 0 0", proj_type="UTM",proj_params="0 0 0 0 0 0 0 0 0 0 0 0 0 0 0",utm_zone=c(10,11),datum="WGS84",pixel_size=250, mosaic = T, proj=T, version='006')

#Show files of some pattern in WD
listfiles <- list.files(pattern='MOD13') # in wd
listfiles<-listfiles[1:5] #or whatever ones are .tif
plot(raster(listfiles[1])) # take a look at what they look like!
#Download, or get shape files separately if you want to crop, my interest here is Oregon.
oregon<-readOGR(dsn="/Users/Maxwell/Documents/geospatial/orcnty2015/", layer = "orcntypoly")
orutm<-spTransform(oregon, crs(raster(listfiles[1]))) #make sure coordinate systems of shapes and MODIS files match
plot(orutm)
```

``` {r}
#make a loop to read in all the tif files and crop them to oregon size, then put them all into a rasterstack
r<-NULL
NDVI<-NULL
for(i in listfiles[1:2]) #just doing two of them for demonstration
  {
 r<-raster(i)/10000 #divide by 10000 to get into correct units.
 #extent(r)<-extent(orutm) #match extents
 r<-crop(r, orutm) #crop to size
 r<-mask(r, orutm) #hide outside pixels
 NDVI<-stack(r, NDVI) #put all modified into a single raster stack
}
```
``` {r, results="markup"}
plot(NDVI)
```